---
title: Estimation of Taylor Rule in the United States in the period 1970:1 - 2020:4
  using the linear regression model
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---
Please, for the context, mathematical and statistical background, database sources read the "README.md" file in the GitHub repository. 

```{r}
rm(list=ls())
```

We can now load the data 
```{r}
intrate <- read.csv("interest_rate_quarterly.csv")
inflrate <- read.csv("inflation_rate_quarterly.csv")
outgap <- read.csv("output_gap_quarterly.csv")
```

Given that our datasets are built in different time periods, we can now solve for time discrepancy 
```{r}
intrate <- intrate[-c(1:22, 227:229), ]; summary(intrate)
inflrate <- inflrate[-c(1:40, 245:248), ]; summary(inflrate)
outgap <- outgap[-c(1:84, 289:292), ]; summary(outgap)
```
We can now rename all columns for clarity purposes
```{r}
names(intrate)[names(intrate) == "Value"] <- "interest_rate"
names(inflrate)[names(inflrate) == "CPALTT01USQ657N"] <- "inflation_rate"
names(outgap)[names(outgap) == "GDPC1_GDPPOT"] <- "output_gap"
```

Create a new database merging the three different sources
```{r}
mydata <- cbind(intrate, inflrate[c("inflation_rate")], outgap[c("output_gap")])
```

We can now create some plots of the original datasets
```{r}
if (!require("ggplot2")) install.packages("ggplot2"); library(ggplot2)
if (!require("reshape2")) install.packages("reshape2"); library(reshape2)

interest_rate <- mydata[["interest_rate"]]
inflation_rate <- mydata[["inflation_rate"]]
output_gap <- mydata[["output_gap"]]
date <- seq(as.Date("1970-01-01"), by="3 month", length.out=204)

dataset <- data.frame(date,interest_rate,inflation_rate,output_gap)
datasetmelt <- melt(dataset, id.vars="date")

ggplot(datasetmelt, aes( x=date, y=value, colour=variable, group=variable )) + 
  geom_line() +
  scale_color_manual(values=c("interest_rate"="black","inflation_rate"="red","output_gap"="orange")) 
```
We can now implement a linear regression with just inflation and output gap as regressors
```{r}
(taylor_reg <- lm(interest_rate ~ inflation_rate + output_gap , data=mydata))
summary(taylor_reg)
```
We can now procede with some tests for linearity, homoskedasticity and uncorrelation. 
```{r}
if (!require("lmtest")) install.packages("lmtest"); library(lmtest)
if (!require("tseries")) install.packages("tseries"); library(tseries)
```
RESET - Ramsey test for linearity
```{r}
resettest(lm(interest_rate ~ inflation_rate + output_gap, data = mydata), power = 2:4, type = "fitted")
```

BREUSCH-PAGAN test for homoskedasticity
```{r}
bptest(taylor_reg)
```
JARQUE-BERA test for normality of the error term
```{r}
residuals <- unlist(taylor_reg[c("residuals")]); typeof(residuals)
jarque.bera.test(residuals)
```
BREUSCH-GODFREY test for uncorrelation
```{r}
bgtest(lm(interest_rate ~ inflation_rate + output_gap, data = mydata), order=3, data=mydata)
```

We implement now a new regression using also unemployment rate as regressors. 
```{r}
unemplrate <- read.csv("unemployment_rate.csv")
unemplrate <- unemplrate[-c(205:208), ]
summary(unemplrate)
names(unemplrate)[names(unemplrate) == "LRUN64TTUSQ156S"] <- "unemployment_rate"

mydata2 <- cbind(mydata, unemplrate[c("unemployment_rate")])

unemployment_rate <- mydata2[["unemployment_rate"]]
dataset2 <- data.frame(date, unemployment_rate, inflation_rate)
dataset2melt <- melt(dataset2, id.vars="date")

colnames(mydata2)
```

We produce now a new plot for the unemployment rate
```{r}
ggplot(dataset2melt, aes( x=date, y=value, colour=variable, group=variable )) + 
  geom_line() +
  scale_color_manual(values=c("inflation_rate"="black","unemployment_rate"="orange")) 
```
Running a linear regression for the new formula
```{r}
(taylor_reg_un <- lm(interest_rate ~ inflation_rate + output_gap + unemployment_rate, data = mydata2))
summary(taylor_reg_un)
```
As before, we procede now running some tests 
RESET - Ramsey test for linearity 
```{r}
resettest(lm(interest_rate ~ inflation_rate + output_gap + unemployment_rate, data = mydata2), power = 2:4, type = "fitted")
```

BREUSCH-PAGAN test for homoskedasticity
```{r}
bptest(taylor_reg_un)
```

JARQUE-BERA test for normality of the error term
```{r}
residuals2 <- unlist(taylor_reg_un[c("residuals")]); typeof(residuals2)
jarque.bera.test(residuals2)
```

BREUSCH-GODFREY test for uncorrelation
```{r}
bgtest(lm(interest_rate ~ inflation_rate + output_gap + unemployment_rate, data = mydata2), order=3, data=mydata)

```


